Formulating RDF Constraints Generically and Validating RDF Data Out-Of-The-Box

### 2015-04-29 paper/paper 2/paper 2.tex
motivation
improved english and little technical change.

%
%
%\title{Formulating RDF Constraints and \\ Validating RDF Data Generically}
\title{Formulating RDF Constraints Generically and Validating RDF Data Out-Of-The-Box }
%\subtitle{}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Erman Acar\inst{2} \and Kai Eckert\inst{2}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{thomas.bosch@gesis.org},\\ 
\and
University of Mannheim, Germany \\
\email{\{erman,kai\}@informatik.uni-mannheim.de} 
}

@@

\textbf{\emph{Complex constraints}} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Exclusive or} is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
The \emph{complex constraint} \emph{context-specific exclusive or of property groups} (\emph{R-13})
restricts individuals of given classes to have properties of only one of multiple mutually exclusive property groups.
\emph{Publications}, e.g., are either identified by an \emph{ISBN} and a \emph{title} (for books) or by an \emph{ISSN} and a \emph{title} (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}, which can be expressed in \emph{ShEx} as follows:

@@

There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
These axiom types are sometimes considered as basic axiom types in \emph{DL} - using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
Such axioms, however, are just syntactic sugar - 
all role characteristics can be expressed using the basic features of \emph{DL}.
The \emph{irreflexive object properties} constraint type (\emph{R-60}) 
restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
With the following \emph{irreflexive object property} constraint, e.g., one can state that individuals cannot be authors of themselves:
\begin{DL}
\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
\end{DL}
When mapped to the \emph{RDF-CV} (see table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):

@@

The \emph{primary key properties} constraint type (\emph{R-226}) is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
\emph{Books}, e.g., are uniquely identified by their \emph{ISBN} number, i.e., the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}, 
which can be represented by the \emph{RDF-CV} in form of a \emph{complex constraint} (see table \ref{tab:primary-key-properties-as-complex-constraints}):

@@

RDF data providers have to ensure that their data conforms to constraints which are explicitly defined within contracts between data providers and consumers.
As there is no standard way to express constraints, they may be represented by a variety of constraint languages\footnote{Candidates for constraint languages are, i.a., SPIN, OWL 2, DSP, ShEx, ReSh, Bibframe, DQTP, Pellet ICV, RDFUnit, SPARQL, and Stardog ICV} - each of them different in syntax and semantics.
To overcome the necessity to understand diverse constraint languages, 
we propose to transform any \emph{specific constraint} (\emph{$sc_{\alpha}$}) (expressed by any constraint language \emph{$\alpha$}) into any other \emph{specific constraint} (\emph{$sc_{\beta}$}) (expressed by any other constraint language \emph{$\beta$}) - in case they have the same semantics.
We use the \emph{RDF-CV} to intermediately represent constraints in a generic way for these transformations.
By defining mappings ($m(sc_{\alpha}), m(sc_{\beta})$) between equivalent \emph{specific constraints} and the corresponding \emph{generic constraint} (\emph{gc}) we are able to convert them automatically:

\begin{DL}
$ gc = m(sc_{\alpha}) = m(sc_{\beta}) $
\end{DL}

We do not need to define mappings between each pair of semantically equivalent \emph{specific constraints}.
Let's assume we are able to express a particular constraint type by 5/6/7/8/9/10 constraint languages.
Without the intermediate \emph{generic constraint}, we would have to define for each constraint type \emph{1} mapping for each pair of \emph{specific constraints} (expressed by different constraint languages)
- that are \emph{\( {n \choose 2} \)} mappings (10/15/21/28/36/45 mappings).
With an intermediate \emph{generic constraint}, however, we only need to define for each constraint type \emph{n} mappings (1 mapping for each \emph{specific constraint}) from \emph{n} \emph{specific constraints} to the corresponding \emph{generic constraint} (5/6/7/8/9/10 mappings).

@@

The \emph{DL} statement \ms{Publication $\equiv$ $\exists$ publisher.Publisher} is mapped to a \emph{(simple) generic constraint} conforming to the \emph{RDF-CV} (see table \ref{tab:existential-quantification-as-simple-constraints}).
Then, one mapping (\emph{$m(sc_{OWL 2})$}, \emph{$m(sc_{ReSh})$}, \emph{$m(sc_{ShEx})$}, ...) is defined for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the corresponding \emph{generic constraint} (\emph{gc}).

@@

We evaluated to which extend the five most promising constraint languages on being the standard fulfill each of the overall 74 requirements to formulate RDF constraints; each of them corresponding to a constraint type.
We recently published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} (serving as appendix of this paper) in which we explain each constraint type in detail and give examples for each (represented by different constraint languages).
The technical report also contains mappings to \emph{DL} to logically underpin each constraint type and to determine which \emph{DL} constructs are needed to express each constraint type \cite{BoschNolleAcarEckert2015}.
If a constraint type can be expressed in \emph{DL}, we added the mapping to \emph{DL} and to the \emph{generic constraint} conforming to the \emph{RDF-CV}.
In contrast, if a constraint type cannot be expressed in \emph{DL}, we only added the mapping to the \emph{generic constraint}.
This way, we prove that each constraint type can be mapped to a \emph{generic constraint} and therefore be represented generically by means of the \emph{RDF-CV}.
Table \ref{tab:evaluation} displays the classification of constraint types according to the dimensions \emph{context}, \emph{complexity}, and \emph{DL expressivity}:

@@

Constraint types can be classified as \emph{property constraints} and \emph{class constraints} (\emph{context}).
Two thirds of the total amount of constraint types are \emph{property constraints} (e.g., \emph{R-86: existential quantification} and \emph{R-91: universal quantification}), one fifth are \emph{class constraints} (e.g., \emph{R-30/37: allowed values} for objects/literals), and approx. 10\% are composed of both \emph{property and class constraints} (e.g., \emph{R-13: context-specific exclusive or of property groups}).
According to the dimesion \emph{complexity}, constraint types can be either atomic (\emph{simple constraints}) or complex (\emph{complex constraints}), i.e., created out of \emph{simple} and/or \emph{complex constraints}.
Almost two thirds of the constraint types are \emph{simple constraints} (e.g., \emph{R-55: object property paths}), a quarter are \emph{complex constraints} (e.g., \emph{R-45: ranges of literal values}), and
nearly 15 percent are \emph{complex constraints} which can be formulated as \emph{simple constraints} when using them in terms of syntactic sugar (e.g., \emph{R-81/75/82/76/80/74: (un)qualified cardinality restrictions}).
Constraint types can either be \emph{expressible in DL} or \emph{not} (\emph{DL expressivity}).
The majority - nearly 70\% - of the overall constraint types are expressible in \emph{DL}.	
\emph{Literal pattern matching} (\emph{R-44}) is an example of a constraint type which cannot be represented in \emph{DL}, but can also be represented generically by means of the \emph{RDF-CV}.
Altogether, most of the constraint types are \emph{simple constraints on properties} which are \emph{expressible in DL}. 
Thus, the majority of constraint types are directly and relatively easy formulated in form of \emph{simple constraints} which can be mapped to equivalent \emph{DL} constructs.

>>>>

%
%
%\title{Formulating RDF Constraints and \\ Validating RDF Data Generically}
\title{Formulating RDF Constraints Generically and Validating RDF Data Out-Of-The-Box }
%\subtitle{}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Erman Acar\inst{2} \and Kai Eckert\inst{3}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS ? Leibniz Institute for the Social Sciences, Germany\\
\email{thomas.bosch@gesis.org},\\ 
\and
University of Mannheim, Germany \\
\email{erman@informatik.uni-mannheim.de} 
\and
Stuttgart Media University, Germany \\
\email{eckert@hdm-stuttgart.de} 
}

@@

\textbf{\emph{Complex constraints}} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Exclusive or} is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
The \emph{complex constraint} \emph{context-specific exclusive or of property groups} (\emph{R-13})
restricts individuals of given classes to have properties of only one of multiple mutually exclusive property groups.
\emph{Publications} are either identified by an \emph{ISBN} and a \emph{title} (for books) or by an \emph{ISSN} and a \emph{title} (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}, which can be expressed in \emph{ShEx} as follows:

@@

There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
These axiom types are sometimes considered as basic axiom types in \emph{DL} - using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
Such axioms, however, are just syntactic sugar - 
all role characteristics can be expressed using the basic features of \emph{DL}.
The \emph{irreflexive object properties} constraint type (\emph{R-60}) 
restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
With the following \emph{irreflexive object property} constraint, for instance, one can state that individuals cannot be authors of themselves:
\begin{DL}
\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
\end{DL}
When mapped to the \emph{RDF-CV} (see table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):

@@

The \emph{primary key properties} constraint type (\emph{R-226}) is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
For instance, \emph{Books} are uniquely identified by their \emph{ISBN} number, that is, the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}, 
which can be represented by the \emph{RDF-CV} in form of a \emph{complex constraint} (see table \ref{tab:primary-key-properties-as-complex-constraints}):

@@

RDF data providers have to ensure that their data conforms to constraints which are explicitly defined within contracts between data providers and consumers.
As there is no standard way to express constraints, they may be represented by a variety of constraint languages\footnote{Candidates for constraint languages are, i.a., SPIN, OWL 2, DSP, ShEx, ReSh, Bibframe, DQTP, Pellet ICV, RDFUnit, SPARQL, and Stardog ICV} - each of them different in syntax and semantics.
To overcome the necessity to understand diverse constraint languages, 
we propose to transform any \emph{specific constraint} (\emph{$sc_{\alpha}$}) (expressed by any constraint language \emph{$\alpha$}) into any other \emph{specific constraint} (\emph{$sc_{\beta}$}) (expressed by any other constraint language \emph{$\beta$}) - in case they have the same semantics.
We use the \emph{RDF-CV} to intermediately represent constraints in a generic way for these transformations.
By defining mappings between equivalent \emph{specific constraints}, and the corresponding \emph{generic constraint} (\emph{gc}) we are able to convert them automatically:

\begin{DL}
$ gc = m_1(sc_{\alpha}) = m_2(sc_{\beta}) $
\end{DL}

We do not need to define mappings between each pair of semantically equivalent \emph{specific constraints}.
Let's assume we are able to express a particular constraint type by 5 constraint languages.
Without the intermediate \emph{generic constraint}, we would have to define for each constraint type \emph{1} mapping for each pair of \emph{specific constraints} (expressed by different constraint languages)
- that are \emph{\( {n \choose 2} \)} mappings which is not very helpful.
With an intermediate \emph{generic constraint}, however, we only need to define for each constraint type \emph{n} mappings (1 mapping for each \emph{specific constraint}) from \emph{n} \emph{specific constraints} to the corresponding \emph{generic constraint}.

@@

The \emph{DL} statement \ms{Publication $\equiv$ $\exists$ publisher.Publisher} is mapped to a \emph{(simple) generic constraint} conforming to the \emph{RDF-CV} (see table \ref{tab:existential-quantification-as-simple-constraints}).
Then, one mapping (\emph{$m_1(sc_{OWL 2})$}, \emph{$m_2(sc_{ReSh})$}, \emph{$m_3(sc_{ShEx})$}, ...) is defined for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the corresponding \emph{generic constraint} (\emph{gc}).

@@

We evaluated to which extend the five most promising constraint languages on being the standard for fulfilling each of the overall 74 requirements to formulate RDF constraints (each corresponds to a constraint type).
We recently published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} (serving as appendix of this paper) in which we explain each constraint type in detail and give examples for each (represented by different constraint languages).
The technical report also contains mappings to \emph{DL} to logically underpin each constraint type and to determine which \emph{DL} constructs are needed to express each constraint type \cite{BoschNolleAcarEckert2015}.
If a constraint type was expressible in \emph{DL}, we added the mapping to \emph{DL} and to the \emph{generic constraint} conforming to the \emph{RDF-CV}.
In contrast, if a constraint type was not expressible in \emph{DL}, we only added the mapping to the \emph{generic constraint}.
This way, we show that each constraint type can be mapped to a \emph{generic constraint} and therefore be represented generically by means of the \emph{RDF-CV}.
Table \ref{tab:evaluation} displays the classification of constraint types according to the dimensions \emph{context}, \emph{complexity}, and \emph{DL expressivity}:

@@

Constraint types can be classified as \emph{property constraints} and \emph{class constraints} (\emph{context}).
Two thirds of the total amount of constraint types are \emph{property constraints} (e.g., \emph{R-86: existential quantification} and \emph{R-91: universal quantification}), one fifth are \emph{class constraints} (e.g., \emph{R-30/37: allowed values} for objects/literals), and approx. 10\% are composed of both \emph{property and class constraints} (e.g., \emph{R-13: context-specific exclusive or of property groups}).
According to the dimesion \emph{complexity}, constraint types can be either atomic (\emph{simple constraints}) or complex (\emph{complex constraints}), i.e., created out of \emph{simple} and/or \emph{complex constraints}.
Almost two thirds of the constraint types are \emph{simple constraints} (e.g., \emph{R-55: object property paths}), a quarter are \emph{complex constraints} (e.g., \emph{R-45: ranges of literal values}), and
nearly 15 percent are \emph{complex constraints} which can be formulated as \emph{simple constraints} when used in terms of syntactic sugar (e.g., \emph{R-81/75/82/76/80/74: (un)qualified cardinality restrictions}).
Constraint types can either be \emph{expressible in DL} or \emph{not} (\emph{DL expressivity}).
The majority - nearly 70\% - of the overall constraint types are expressible in \emph{DL}.	
\emph{Literal pattern matching} (\emph{R-44}) is an example of a constraint type which cannot be represented in \emph{DL}, but can also be represented generically by means of the \emph{RDF-CV}.
Altogether, most of the constraint types are \emph{simple constraints on properties} which are \emph{expressible in DL}. 
Thus, the majority of constraint types are directly and relatively easy formulated in form of \emph{simple constraints} which can be mapped to equivalent \emph{DL} constructs.
